### 2012-07-09 Doc/VRIPHYS2012/vriphys2012.tex
changing SCA 2012 to Vriphys 2012
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@419 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

In the rest of the document we define the surface over the triangle:
\begin{equation}
    (\xi_1,\xi_2) \in \Delta^2 = \left\{ (\xi_1,\xi_2)~|~\xi_1, \xi_2 \ge 0
        \mathrm{~and~} \xi_1+\xi_2 \le 1 \right\}
\end{equation}

@@

\begin{enumerate}
    \item The plane perpendicular to the normal at the vertex.
    \item The plane that contains the the curve of triangle's contour. The
        choise is arbitrary, but necessary to maintain $C^0$ continuity. 
        We choose the plane defined by the edge of the flat triangle and
        average of the two normals at vertices of the edge. 
    \item The plane perpendicular to the edge of the flat triangle placed at
        $1/3$ of the edge length.
\end{enumerate}

@@

$C^1$ continuity can also be maintained if special care is taken when
computing the position of the central point (see \cite{Ubach2010}). For
simplicity we compute the central point followingly:

@@

We base the displacement field $u$ on the formulation of the Bézier
triangle \eqref{eq-cubicbez}:

@@

\begin{equation}
    P_{i-j} = P_j-P_i, \quad i \in [4;10], j \in [1;3]
\end{equation}

@@

\begin{equation}
    J_m = \left[ \begin{matrix}
        \deriv{u_x}{x} \\
        \deriv{u_y}{y} \\
        \deriv{u_x}{y} + \deriv{u_y}{x}
    \end{matrix} \right]
\end{equation}

@@

\begin{equation}
    J_b = \left[ \begin{matrix}
        - z \deriv{^2 u_z}{x^2} \\
        - z \deriv{^2 u_z}{y^2} \\
        - 2z \deriv{^2 u_z}{xy}
    \end{matrix} \right]
\end{equation}

@@

where $M$ is the material matrix. We use Hooke's law.

Because the deformation field for the shell is non-linear \eqref{eq-bezU}
the integrals \eqref{eq-Km} and \eqref{eq-Kb} have to be computed using
numerical integration. In our implementation we employ 4-point Gaussian
quadrature for integration over triangle area.
% TODO: add ref
% TODO: add abscissas/weights?
%  \xi_1        \xi_2           w
% 0.211324865, 0.166666667 0.197168783
% 0.211324865, 0.622008467 0.197168783
% 0.788675134, 0.044658198 0.052831216
% 0.788675134, 0.166666667 0.052831216

To keep the system linear during simulation we compute the forces in
corotational frame.

@@

\section{Mechanical Mapping} % {{{

% TODO: mention it is 6 DOF <-> 3 DOF

Because of the bending property of the shells relatively few elements are
necessary to simulate curved surface. To enrich the visual experience from
the simulated object it is desirable to use more triangles in the areas
with high curvature during rendering. To do that we can map a high
resolution mesh onto the mechanical mesh. For every vertex of the high
resolution mesh we first find the triangle on the mechanical mesh that is
closest to the vertex and then assign barycentric coordinates on the
triangle to this vertex. After every step of the simulation the high
resolution mesh is updated using the assigned barycentric coordinates and
the function of the surface \eqref{eq-cubicbez}.

mapping velocities (necessary for collisions?)

@@

Which means we use the same expression to interpolate the velocities only
by substituting the velocities at control points. For the corner nodes we
already have the velocities $V_i$. For the internal nodes we need to compute
them. In case of the edge nodes it is a problem of computing the velocity
of a point attached to the moving rigid body:

@@

* reverse mapping forces for collision response

We compute the influence of force $F$ on the corners through all the
control points by the formula:

\begin{equation}
    F_j += F n_i
\end{equation}

and the torques applied through edge control points by:

\begin{eqnarray}
    \theta_j += (P_i - P_j) \times F
\end{eqnarray}

where $j$ is the index of the corner node associated with control point
$P_i$.

And finaly we add the torques through the central control point with:

\begin{eqnarray}
    \theta_1 += ((P_4 - P_1) + (P_5 - P_1)) \times F \\
    \theta_2 += ((P_6 - P_2) + (P_7 - P_2)) \times F \\
    \theta_3 += ((P_8 - P_3) + (P_9 - P_3)) \times F
\end{eqnarray}

>>>>

In the rest of the document we define the surface over the triangle:
\begin{equation}
    (\xi_1,\xi_2) \in \bigtriangleup = \left\{ (\xi_1,\xi_2)~|~\xi_1, \xi_2 \ge 0
        \mathrm{~and~} \xi_1+\xi_2 \le 1 \right\}
\end{equation}

@@

\begin{enumerate}
    \item The plane perpendicular to the normal at the vertex.
    \item The plane that contains the curve of triangle's contour. The
        choise is arbitrary, but necessary to maintain $C^0$ continuity. 
        We choose the plane defined by the edge of the flat triangle and
        average of the two normals at vertices of the edge. 
    \item The plane perpendicular to the edge of the flat triangle placed at
        $1/3$ of the edge length.
\end{enumerate}

@@

$C^1$ continuity on the edges can also be maintained if special care is taken
when computing the position of the central point (see \cite{Ubach2010}). For
simplicity we compute the central point followingly:

@@

Our definition of the displacement field $u$ is based on the formulation of the
Bézier triangle \eqref{eq-cubicbez}:

@@

\begin{equation}
    P_{i-j} = P_j-P_i %, \quad i \in [4;10], j \in [1;3]
\end{equation}

@@

\begin{equation}\label{eq-cauchy}
    J_m = \left[ \begin{matrix}
        \deriv{u_x}{x} \\
        \deriv{u_y}{y} \\
        \deriv{u_x}{y} + \deriv{u_y}{x}
    \end{matrix} \right]
\end{equation}

@@

\begin{equation}\label{eq-kirchhoff}
    J_b = \left[ \begin{matrix}
        - z \deriv{^2 u_z}{x^2} \\
        - z \deriv{^2 u_z}{y^2} \\
        - 2z \deriv{^2 u_z}{xy}
    \end{matrix} \right]
\end{equation}

@@

where $M$ is the material matrix. To keep the system simple we use linear
Hooke's law for isotropic materials.

Because the deformation field for the shell is non-linear \eqref{eq-bezU}
the integrals \eqref{eq-Km} and \eqref{eq-Kb} have to be computed using
numerical integration. In our implementation we employ 6-point Gaussian
quadrature for integration over triangle area.
% TODO: add ref
% TODO: add abscissas/weights?

The Cauchy's strain tensor \eqref{eq-cauchy} is know to not be rotation
invariant \footnote{add ref}. To deal with rigid body rotations we compute
the displacements and forces in corotational frame. At each step of the
simulation we compute the local frame for every element from the position
of the corner nodes and we allign one axis of the frame with one edge of
the element. \footnote{add image}

The resulting internal force of deformed elemnt is then computed relative
to this frame:

\begin{equation}
    F_e = R_e^T (K_e u) = R_e^T \left(K_e(R_e x - R_e^0 x^0)\right)
\end{equation}

\noindent
where $R$ and $R_0$ are rotation matrices of the frame for current and rest
mesh respectively.

@@

\section{Visual and Mechanical Mapping} % {{{

Because of the bending property of the shells relatively few elements are
necessary to simulate curved surface. To enrich the visual experience from
the simulated object it is desirable to use more triangles in the areas
with high curvature during rendering. Besides the visual accuracy the high
resolution mesh correctly modeling the curved areas is needed for
interaction and correct simulation of collisions and constraints. To handle
these we have also to know how to map the velocities from the mechanical
mesh onto the high resolution mesh and how to map forces acting on the high
resolution mesh back onto the mechanical mesh.

\subsection{Mapping the Surface Mesh}

We use the fact that the geometry of the shell is based on the formulation
of cubic Bézier triangle \eqref{eq-cubicbez}. For every vertex of the high
resolution mesh we first find the corresponding triangle on the mechanical
mesh that is closest to the vertex and assign barycentric coordinates on
the triangle to this vertex. After every step of the simulation the high
resolution mesh is updated using the assigned barycentric coordinates and
the function of the surface \eqref{eq-cubicbez}.

\subsection{Mapping Velocities}

In the

@@

Which means we use the same expression to interpolate the velocities only
substituting the velocities at control points. For the corner nodes we
already have the velocities $V_i$. For the internal nodes we need to compute
them. In case of the edge nodes it is a problem of computing the velocity
of a point attached to the moving rigid body:

@@

\subsection{Mapping Forces}

For physical interaction with the object one also requiers that the normal
forces applied on the high resolution mesh are trasmitted back as forces
and torques on the vertices of the mechanical mesh.

We compute the influence of force $F$ on the corners through all the
associated control points by the formula:

\begin{eqnarray}
    F_1 & = & F \sum_{i \in \{1,4,5\}} n_i + \frac{1}{3} F n_{10} \\
    F_2 & = & F \sum_{i \in \{2,6,7\}} n_i + \frac{1}{3} F n_{10} \\
    F_3 & = & F \sum_{i \in \{3,8,9\}} n_i + \frac{1}{3} F n_{10} \\
\end{eqnarray}

%\begin{eqnarray}
%    F_1 & = & F (n_1 + n_4 + n_5) + \frac{1}{3} F n_{10} \\
%    F_2 & = & F (n_2 + n_6 + n_7) + \frac{1}{3} F n_{10} \\
%    F_3 & = & F (n_3 + n_8 + n_9) + \frac{1}{3} F n_{10} \\
%\end{eqnarray}

Similarily we compute the torques applied through edge control points and
the central control points by:

\begin{eqnarray}
    \theta_1 & = & P_{1-4} \times (n_4 F) + P_{1-5} \times (n_5 F) + \\
        & + & \frac{1}{3} (P_{1-4} + P_{1-5}) \times (n_{10} F) \\
    \theta_2 & = & P_{1-6} \times (n_6 F) + P_{1-7} \times (n_7 F) + \\
        & + & \frac{1}{3} (P_{1-6} + P_{1-7}) \times (n_{10} F) \\
    \theta_3 & = & P_{1-8} \times (n_8 F) + P_{1-9} \times (n_9 F) + \\
        & + & \frac{1}{3} (P_{1-8} + P_{1-9}) \times (n_{10} F) \\
\end{eqnarray}

%\begin{eqnarray}
%    \theta_1 & = & \sum_{i \in \{4,5\}} P_{1-i} \times (n_i F) \\
%    & + &\frac{1}{3} (P_{1-4} + P_{1-5}) \times (n_{10} F) \\
%    \theta_1 & = & \sum_{i \in \{6,7\}} P_{1-i} \times (n_i F) + \\
%    & + &\frac{1}{3} (P_{1-6} + P_{1-7}) \times (n_{10} F) \\
%    \theta_1 & = & \sum_{i \in \{8,9\}} P_{1-i} \times (n_i F) + \\
%    & + &\frac{1}{3} (P_{1-8} + P_{1-9}) \times (n_{10} F) \\
%\end{eqnarray}
%
%\begin{eqnarray}
%    \theta_1 & = & \sum_{i \in \{4,5\}} P_{1-i} \times (n_i F) +
%    \frac{1}{3} (P_{1-4} + P_{1-5}) \times (n_{10} F) \\
%    \theta_1 & = & \sum_{i \in \{6,7\}} P_{1-i} \times (n_i F) +
%    \frac{1}{3} (P_{1-6} + P_{1-7}) \times (n_{10} F) \\
%    \theta_1 & = & \sum_{i \in \{8,9\}} P_{1-i} \times (n_i F) +
%    \frac{1}{3} (P_{1-8} + P_{1-9}) \times (n_{10} F) \\
%\end{eqnarray}
